### 2017-07-17 lafaySimulatedScenePerception.tex
ml proof read 4

///

@@

First, in order to ensure the coherence of the data, we check that none of the ni- scenes gets a $\mathcal{A}_{scene}$ higher than one of a i- scene. Four ni- scenes don’t respect that rule: they are removed, together with their corresponding i- scenes. 36 i -scenes and 36 ni -scenes remain for analysis. Secondly, we want to be sure that participants really perceived a difference in terms of pleasantness between i- and ni -scenes. For that, we investigate the mean pleasantness score for each participant $\mathcal{A}_{subject}$, computed separately for each type of environment. It appears that the i- scenes were perceived significatively more pleasant ($p < 0.01$) than the ni- scenes.

@@

First, we focus on the sound level. Figures ~\ref{fig:soundlevela},~\ref{fig:soundlevelb} and~\ref{fig:soundlevelc} depict the levels $L$, $L(E)$ and $L(T)$. There is a significative difference in terms of sound level between i- and ni -scenes ($L$: $p<0.01$), with a standard deviation of -7 $dB$. This difference is siginificant for events ($L(E)$: $p<0.01$, STD: -7 $dB$) and textures ($L(T)$: $p<0.01$, STD: -6 $dB$).

As expected, we verify that the sound sources sound level is indeed a pleasantness indicator, the ni -scenes tending to be louder - what also comes out from a large number of related works. We also notice that this difference of sound level is equally true for events or textures.

It appears that the biggest influence on the global sound level comes from the events, the gap between $L$ and $L(E)$ being only 1 $dB$ for i- and ni- scenes. This observation is in agreement with the results obtained by Kuwano~\al~\cite{kuwano_memory_2003}. During their experiment, the authors ask to their participants to assess a set of soundscapes at a global level and then to do the same judgment at the moments when each of them detect a sound source. The study shows that there is no significative difference between global and averaged instantaneous judgments. In our case, the result can be interpreted as if the participants had unconsciously integrated this perceptual reality when composing the scenes, by assigning the global sound level to well identified short sounds, \ie the events.

The level alone is not enough to differentiate the different types of environment. In fact, $20\%$ of the i- scenes hast got a sound level higher than the nominal level of the ni- scenes, while there is no overlap when considering the perceived pleasantness $\mathcal{A}_{scene}$.

@@

In this section, relationships that could exist between structural features and perceived pleasantness are investigated. Contrary to the previous section, we do not consider the emotional property of a scene on a binary mode (i vs. ni): here, we consider the mean pleasantness $\mathcal{A}_{scene}$ as a perceptual feature. The aim is to investigate the level of correlation between structural features and $\mathcal{A}_{scene}$. The linear correlation coefficients computed between $\mathcal{A}_{scene}$ \vs~$L$, $L(E)$, $L(T)$ are shown on Table~\ref{tab:corrStructA}. Relationships between $\mathcal{A}_{scene}$ and the structural features are depicted in Figure~\ref{fig:soundleveld},~\ref{fig:soundlevele} and ~\ref{fig:soundlevelf}.

@@

When considering the whole set of scenes, this result can be explained by the fact the i- scenes tends to be softer than the ni- scenes, deluding that the negative correlation observed for the ni- scenes is extended.

@@

\begin{itemize}
\item allows to differentiate between i- and ni- scenes,
\item allows a fine characterization of the perceived pleasantness for ni- scenes,
\item is not a relevant feature for the perceived pleasantness for a priori pleasant soundscapes (i- scene)
\end{itemize}

The same conclusions can be drawn for $L(E)$ (\cf~Figure~\ref{fig:soundlevele}). For $L(T)$ (\cf~Figure~\ref{fig:soundlevelf}), the moderate correlation observed for the whole set of scenes disappears when separate scenes are considered (i-scenes: $r=-0.33$, $p=0.05$, ni-scenes: $r=-0.00$, $p=0.99$). One can suppose that the negative correlation coming from the whole is an artefact due to the level difference between the two types of scenes (i- scenes tend to be softer than ni - scenes). Thus, the sound events keep a relative ability to predict the pleasantness of the ni- scenes, but the level of the textures do not bring a lot of information whatever the type of environment is.

@@

Those first outcomes tend to show that two modes of perception exists depending on the nature of the environment (i- or ni-), and each involve independent features. The fact that $L$ is not able to characterize the pleasantness of the i- scenes can lead us to think that all the sound sources do not equally contribute to the perception of pleasantness, but that only the level of some of them can influence this percept. In order to go deeper in that direction, in the next section, we analyze these scenes on a semantic point of view, \ie we take an interest in the nature of the sources they are made of.

@@

We analyse the composition of the scenes by counting the number of participants who used a given class of sounds to simulate the environment. Results are shown on Figure~\ref{fig:soundsourcea} for events and on Figure~\ref{fig:soundsourceb} for textures. For sake of clarity, we chose a transitional level of abstraction between level 0 and 1, named $0+$, to depict classes, see ~Figure~\ref{fig:taxonomie}.

We observe a significant difference between i- and ni- scenes for the classes choice. The distribution of the classes is very similar to to the one obtained in a related work on ideal urban soundscapes \cite{guastavino2006ideal}, \ie~lasses involving human presence and nature are prevailing in the i- scenes, and on the opposite, classes involving mechanical sounds and/or public works are prevailing in the ni- scenes. These results confirm a previously observed fact: the semantic nature of the sound sources play an important role in the assessment of the environment \cite{raimbault2005urban,dubois2006cognitive}.

Nevertheless, we notice some differences with \cite{guastavino2006ideal}: Guastavino’s results show that sounds of \emph{public transportation} are specific of ideal urban soundscapes. The authors interpret this by the fact that the perception of pleasantness is, among other things, due to socio-cultural factors. In our representation of the world, sounds of public transportation are positively connoted and tend to be more accepted than sounds of personal vehicles.

To a certain extent, our results are in contradiction with this assumption. In fact,  Figure~\ref{fig:soundsourcea} show \emph{public transportation} classes (\emph{bus} and \emph{train}, \cf~Figure~\ref{fig:soundsourcec}) have been used by the participants for $28\%$ of the i-scenes, and for $42\%$ of the ni-scenes. the results do not re-evaluate the fact that sounds of \emph{public transportation} are well accepted: $25\%$ of the subjects used \emph{bus} pour les i-scenes, a level that is comparable to the of the \emph{bike} class, and much higher than all the personal vehicles classes. Though, \emph{public transport} classes are also strongly present in the ni-scenes, for instance more than \emph{light vehicle} or \emph{trucks} classes. On the basis of our results, the public transportation class cannot be considered as typical of an ideal urban soundscape.

This difference can be explained by the nature of the experimental protocol used in the two studies. As in our case, Guastavino asks the participants for describing an environment; but in that case, they work from their memories whereas, in our case, participants do that task using sound samples. The fact that participants in our experiment are faced to the acoustic reality of the sounds for composing the environment may have decreased the socio-cultural impact. Other works using sounds as stimuli have shown that the \emph{bus} class can have a negative influence on the assessment of the environment \cite{lavandier2006contribution}.

@@

We highlight that, on a qualitative point of view, the composition of the scenes in terms of sound sources differs regarding the type of environment (i or ni). We are now trying to to investigate if some of the sound classes are specific to a given environment. For that, we use the V-test as detailed in  Section~\ref{sec:xp1_dataAna}, while considering separately each abstraction level. Results are presented in Table~\ref{tab:markers}.

Regarding the sound events, 9 markers are identified at all the abstraction levels. As Figure~\ref{fig:soundsource} could let it predict, classes related to human presence (\emph{male footsteps on concrete}, \emph{bicycle bell}), and of nature (\emph{animals, bird}, and \emph{bird song}) are i- scenes markers; we also notice the \emph{church bell} class as an i- scenes marker. This latter result may be due to the socio-cultural background of the participants who are mostly European citizens. In fact, according to Schafer, a sound that is identified by a person as being an element of his/her environment, is well accepted. Sound markers of ni- scenes are classes related to construction site (\emph{construction works}), or suggesting a strong traffic (\emph{horn}, \emph{siren}).

Regarding the textures, 5 markers are identified. For the i- scenes, they are classes related to passive or quiet ambiances (\emph{courtyard}, \emph{park}). for the ni- scenes, as for the events, they are classes related to construction site (\emph{construction}, \emph{construction vehicle}), together with a class related to traffic (\emph{crossroad}).

Although the whole set of identified markers are intuitive, none of the event classes related to the noise of the motor vehicles is a marker, except for the texture class \emph{crossroad}. To simulate an unpleasant traffic, participants chose the classes \emph{horn} or \emph{siren}. One can suppose that isolated motor vehicle sounds are understood as taking fully part of the urban environment, and thus are not especially linked to an unpleasant soundscape.

@@

In this section, we evaluate the ability of a semantic representation to split the two types of environment. For that, we compute a rank 5-precision ($p@5$) on the space induced by the semantic features $S$, and for each abstraction level, see Section~\ref{sec:xp1_dataAna}. The vectors $S$ are built by using all the classes ($ET$), only the event classes ($E$), only the texture classes ($T$), only the event classes corresponding to sound markers ($E_m$), or only the event classes excluding sound markers ($E_{w/o,m}$). Texture classes corresponding to sound markers are not considered as these latter are not numerous enough. For the same reasons, event classes corresponding to sound markers at abstraction level $0$ are also discarded. Results are shown on Figure~\ref{fig:pa5}.

Concerning $ET$, the $p@5$ is $76\%$ for the abstraction level $0$, and stay above $86\%$ from the abstraction level $1$. These results confirm that it is possible to make a clear distinction between the two types of environment on the unique basis of presence / absence of sound classes. We also notice that the higher the abstraction level is, the more important is the ability to make this distinction. In other words, the more one is precise in the scene description, the more one is able to clearly differentiate between i- and ni- scenes.

Considering $E$ and $T$ separately, ir appears: 1) that the $p@5$ obtained with $E$ is similar to the one obtained with $ET$ ; 2) that the $p@5$ obtained with $T$ is systematically lower than the one obtained with $E$, by $10\%$ or $15\%$. Those results indicate that the semantic information allowing to split the two environments is mainly contained in the events. Moreover, these results are in line with the work done in \cite{maffiolo_caracterisation_1999} showing that we analyze in a causal manner (i.e. by identifying the sources) the event scenes which are composed of several sound events, see Section~\ref{sec:simscene_sampleDataSet}.

>>>>

@@

First, in order to ensure the coherence of the data, we check that none of the ni- scenes gets a $\mathcal{A}_{scene}$ higher than one of a i- scene. Four ni- scenes don’t fullfill that constraint: they are thus removed, together with their corresponding i- scenes. As a consequence, 36 i -scenes and 36 ni -scenes remain for analysis. Second, we verify that participants really perceived a difference in terms of pleasantness between i- and ni -scenes. For that, we investigate the mean pleasantness score for each participant $\mathcal{A}_{subject}$, computed separately for each type of environment. It appears that the i- scenes were perceived significatively more pleasant ($p < 0.01$) than the ni- scenes.

@@

First, our analyses focuses on the sound level. Figures ~\ref{fig:soundlevela},~\ref{fig:soundlevelb} and~\ref{fig:soundlevelc} respectively depict the levels $L$, $L(E)$ and $L(T)$. There is a significative difference in terms of sound level between i- and ni -scenes ($L$: $p<0.01$), with a standard deviation of -7 $dB$. This difference is siginificant for events ($L(E)$: $p<0.01$, STD: -7 $dB$) and for textures ($L(T)$: $p<0.01$, STD: -6 $dB$).

As expected, the sound sources sound level is indeed a pleasantness indicator, as the ni -scenes tend to be louder, an outcome of a large number of related works. We also notice that this difference of sound level is equally true for events or textures.

It appears that the biggest influence on the global sound level comes from the events, the gap between $L$ and $L(E)$ being only 1 $dB$ for i- and ni- scenes. This observation is in agreement with the results obtained by Kuwano~\al~\cite{kuwano_memory_2003}. During their experiment, the authors ask to their participants to assess a set of soundscapes at a global level and then to do the same judgment at the time when they detect a sound source. The study shows that there is no significative difference between global and averaged instantaneous judgments. In our case, the result can be interpreted as if the participants had unconsciously integrated this perceptual reality when composing the scenes, by assigning the global sound level to well identified short sounds, \ie the events.

Though, the level alone is not sufficient to differentiate the type of environment. In fact, $20\%$ of the i- scenes have a sound level higher than the nominal level of the ni- scenes, while there is no overlap when considering the perceived pleasantness $\mathcal{A}_{scene}$.

@@

In this section, relationships that could exist between structural features and perceived pleasantness are investigated. Contrary to the previous section, we do not consider the pleasantness property of a scene on a binary mode (i vs. ni): this is the mean pleasantness $\mathcal{A}_{scene}$ that is considered here as a perceptual feature. The aim is to investigate the level of correlation between structural features and $\mathcal{A}_{scene}$. The linear correlation coefficients computed between $\mathcal{A}_{scene}$ \vs~$L$, $L(E)$, $L(T)$ are shown on Table~\ref{tab:corrStructA}. Relationships between $\mathcal{A}_{scene}$ and the structural features are depicted in Figure~\ref{fig:soundleveld},~\ref{fig:soundlevele} and ~\ref{fig:soundlevelf}.

@@

When considering the whole set of scenes, the fact that the level is indeed a good indicator of pleasantness can be explained by the fact that the i- scenes tend to be softer than the ni- scenes, thus allowing us to extend to the i-scenes the negative correlation observed for the ni- scenes.

@@

\begin{itemize}
\item allows us to differentiate between i- and ni- scenes,
\item characterizes precisely the perceived pleasantness for ni- scenes,
\item is not a relevant feature for modeling the perceived pleasantness of \textit{a priori} pleasant soundscapes (i- scenes)
\end{itemize}

The same conclusions can be drawn for $L(E)$, see~Figure~\ref{fig:soundlevele}. For $L(T)$, as shown on~Figure~\ref{fig:soundlevelf}, the moderate correlation observed for the whole set of scenes disappears when separate scenes are considered (i-scenes: $r=-0.33$, $p=0.05$, ni-scenes: $r=-0.00$, $p=0.99$). Again, we believe that the negative correlation coming from the whole is an artefact due to the level difference between the two types of scenes (i- scenes tend to be softer than ni - scenes). Thus, the sound events keep a relative ability to predict the pleasantness of the ni- scenes, but the level of the textures do not bring a lot of information whatever the type of environment is.

@@

Those first outcomes tend to show that 1) two modes of perception exists depending on the very nature of the environment (i- or ni-), and 2) each involve independent features. The fact that $L$ is not sufficient to characterize the pleasantness of the i- scenes can lead us to conclude that all the sound sources do not equally contribute to the perception of pleasantness. Thus, we think that only the level of some of them can influence this percept. In order to investigate further in that direction, we analyze in the next section the scenes on a semantic point of view, \ie we take an interest in the nature of the sources they are made of.

@@

We study the composition of the scenes by counting the number of participants who used a given class of sounds to simulate the environment. Results are shown on Figure~\ref{fig:soundsourcea} for events and on Figure~\ref{fig:soundsourceb} for textures. For the sake of clarity, a transitional level of abstraction between level 0 and 1, named $0+$ is used to depict classes, see ~Figure~\ref{fig:taxonomie}.

We observe a significant difference between i- and ni- scenes for the classes choice. The distribution of the classes is very similar to the one obtained in a related work on ideal urban soundscapes \cite{guastavino2006ideal}, \ie~classes involving human presence and nature are prevailing in the i- scenes, and on the opposite, classes involving mechanical sounds and/or public works are prevailing in the ni- scenes. These results confirm a previously observed fact: the semantic nature of the sound sources play an important role in the assessment of the environment \cite{raimbault2005urban,dubois2006cognitive}.

Nevertheless, we notice some differences with \cite{guastavino2006ideal}: Guastavino’s results show that sounds of \emph{public transportation} are specific of ideal urban soundscapes. The authors interpret this by the fact that the perception of pleasantness is, among other things, due to socio-cultural factors. Thus, in our representation of the world, sounds of public transportation are positively connoted and tend to be more accepted than sounds of personal vehicles.

To a certain extent, our results are in contradiction with this assumption. In fact,  Figure~\ref{fig:soundsourcea} show that \emph{public transportation} classes (\emph{bus} and \emph{train}, \cf~Figure~\ref{fig:soundsourcec}) have been used by the participants for $28\%$ of the i-scenes, and for $42\%$ of the ni-scenes. the results do not re-evaluate the fact that sounds of \emph{public transportation} are well accepted: $25\%$ of the subjects used \emph{bus} pour les i-scenes, a level that is comparable to the of the \emph{bike} class, and much higher than all the personal vehicles classes. Though, \emph{public transport} classes are also strongly present in the ni-scenes, for instance more than \emph{light vehicle} or \emph{trucks} classes. On the basis of our results, the public transportation class cannot be considered as typical of an ideal urban soundscape.

This difference may be explained by the nature of the experimental protocol used in the two studies. As in our study, Guastavino asks the participants for describing an environment starting from their memories whereas, in our case, participants perform the same task using actual sound samples. The fact that participants in our experiment are faced to the acoustic reality of the sounds for composing the environment may have decreased the socio-cultural impact. Other works using sounds as stimuli have shown that the \emph{bus} class can have a negative influence on the assessment of the environment \cite{lavandier2006contribution}.

@@

We highlight that, on a qualitative point of view, the composition of the scenes in terms of sound sources differs with respect to the type of environment (i or ni). Let us now investigate whether some of the sound classes are specific to a given environment. For that, we use the V-test as detailed in  Section~\ref{sec:xp1_dataAna} separately for each abstraction level. Results are presented in Table~\ref{tab:markers}.

Regarding the sound events, 9 markers are identified for all abstraction levels. As shown on Figure~\ref{fig:soundsource}, classes related to human presence (\emph{male footsteps on concrete}, \emph{bicycle bell}), and of nature (\emph{animals, bird}, and \emph{bird song}) are i- scenes markers as well as the \emph{church bell} class. This latter result may be due to the socio-cultural background of the participants who are mostly European citizens. In fact, according to Schafer, a sound that is identified by a person as being an element of his/her environment, is well accepted. Sound markers of ni- scenes are classes related to construction site (\emph{construction works}), or suggesting a strong traffic (\emph{horn}, \emph{siren}).

Regarding the textures, 5 markers are identified. For the i- scenes, they are classes related to passive or quiet ambiances (\emph{courtyard}, \emph{park}). For the ni- scenes, as for the events, they are classes related to construction site (\emph{construction}, \emph{construction vehicle}), together with a class related to traffic (\emph{crossroad}).

Although the whole set of identified markers are rather intuitive, none of the event classes related to the noise of the motor vehicles are identified as markers, except for the texture class \emph{crossroad}. To simulate an unpleasant traffic, participants chose the classes \emph{horn} or \emph{siren}. We thus conclude that isolated motor vehicle sounds are understood as being part of the urban environment, and thus their nature is not necessarily linked to an unpleasant soundscape.

@@

In this section, we evaluate the ability of a semantic representation of the scenes to discriminate between the two types of environment. For that, we consider a rank 5-precision ($p@5$) on the space induced by the semantic features $S$, and for each abstraction level, see Section~\ref{sec:xp1_dataAna}. The vectors $S$ are built by using all the classes ($ET$), only the event classes ($E$), only the texture classes ($T$), only the event classes corresponding to sound markers ($E_m$), or only the event classes excluding sound markers ($E_{w/o,m}$). Texture classes corresponding to sound markers are not considered as these latter are too few to reliably compute the metric. For the same reasons, event classes corresponding to sound markers at abstraction level $0$ are also discarded. Results are shown on Figure~\ref{fig:pa5}.

Concerning $ET$, the $p@5$ is $76\%$ for the abstraction level $0$, and stay above $86\%$ from the abstraction level $1$. These results confirm that it is possible to make a clear distinction between the two types of environment on the sole basis of presence / absence of sound classes. We also notice that the higher the abstraction level is, the more important the ability to make this distinction is. In other words, the more precise in the scene description is, the more clear the discrimination between i- and ni- scenes is.

Considering $E$ and $T$ separately, it appears that: 1) the $p@5$ obtained with $E$ is similar to the one obtained with $ET$ ; 2) the $p@5$ obtained with $T$ is systematically lower than the one obtained with $E$, by $10\%$ or $15\%$. Those results indicate that the discriminative semantic information is mainly contained in the events. These results are in line with the work done in \cite{maffiolo_caracterisation_1999} showing that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie by identifying the sources, as discussed in Section~\ref{sec:simscene_sampleDataSet}.
